---
title: "Exploratory Data Analysis - Single bakery item"
output: github_document

knit: (function(inputFile, encoding) {
      out_dir <- "../markdowns";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(here)

# Read in data
train <- read_csv(here("data/train.csv")) %>%  
  filter(item_nbr == 502331)
test <- read_csv(here("data/test.csv")) %>% 
  filter(item_nbr == 502331)
items <- read_csv(here("data/items.csv")) %>% 
  filter(family == "BREAD/BAKERY")
stores <- read_csv(here("data/stores.csv"))
transactions <- read_csv(here("data/transactions.csv"))
oil <- read_csv(here("data/oil.csv"))
holidays <- read_csv(here("data/holidays_events.csv"))
```

## Training data

### Date continuity and missing values

* The 359th (360th on a leap year) day is Christmas day and there is no data for this day
  
```{r}
train %>% 
  mutate(day = lubridate::yday(date)) %>% 
  group_by(year = year(date)) %>% 
  summarise(n = n(),
            min = min(day),
            max = max(day),
            distinct_days = n_distinct(day),
            NAs = sum(is.na(day)),
            missing_days = paste(setdiff(1:366,day), collapse = ", "))
```

### Visualise time series

Monthly sales of bakery/bread item in each store location
    
```{r}
## Month 
train %>% 
  mutate(dom = ymd(paste(year(date),month(date),1, sep = "-"))) %>% 
  count(dom, store_nbr, wt = unit_sales) %>% 
  ggplot(aes(x = dom, y = n, col = store_nbr, group = store_nbr)) +
  geom_line() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 month") +
  labs(x = "Date",
       y = "Monthly sales") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```


### Daily sales by store


* Some stores have large outliers (29, 33, 34)
* The median sales are relatively similar
```{r}
train %>% 
  mutate(store_nbr = factor(store_nbr),
         store_nbr = fct_reorder(store_nbr, unit_sales, median)) %>% 
  ggplot(aes(x = unit_sales, y = store_nbr)) +
  geom_boxplot()
```

### Total sales by day of the week

* Sunday has the largest volume of sales, significantly more than other days
* Saturday has the second largest volume of sales
* There is a smaller variance between total sales for other days of the week

```{r}
train %>% 
  arrange(store_nbr, date) %>% 
  mutate(dow = lubridate::wday(date, label = TRUE)) %>% 
  group_by(store_nbr, dow) %>% 
  summarise(unit_sales_total = sum(unit_sales),
            n = n(),
            .groups = "drop") %>% 
  mutate(store_nbr = factor(store_nbr),
         store_nbr = fct_reorder(store_nbr, unit_sales_total, sum)) %>% 
  ggplot(aes(x = unit_sales_total, y = store_nbr, col = dow)) + 
  geom_point() +
  ggrepel::geom_label_repel(data = ~group_by(., dow) %>% sample_n(size = 1),
                            aes(label = dow),
                            min.segment.length = 0,nudge_x = -5000)
```

### Daily sale box plot

A sample of 15 stores

* Clearly shows as above, Sunday and Saturday are the bigest sale days

```{r, fig.height=11}
train %>% 
  filter(store_nbr %in% sample(1:54,15)) %>% 
  arrange(store_nbr, date) %>% 
  mutate(dow = lubridate::wday(date, label = TRUE, week_start = 1),
         store_nbr = factor(store_nbr),
         store_nbr = fct_reorder(store_nbr, unit_sales, sum)) %>% 
  ggplot(aes(x = unit_sales, y = store_nbr, fill = dow)) + 
  geom_boxplot()
```
